### 2018-01-27 ApproxComp/sec-data.tex

Link prediction methods are often applied to very large and sparse networks, which have a large {\em search space} $O(n^2)$,
where $n$ is the number of nodes. Hence, the scalability is a big challenge. In fact, an often overlooked fact is that most {\em exiting link prediction algorithms evaluate the link propensities only over a subset of possibilities rather than all propensities over the entire network},
\eg \cite{zhu2016}.
In order to understand why this is the case, consider a large network with $10^8$ nodes. Its number of {\em possibilities} for links
is of the order of $10^{16}$. Therefore, a 3GHz processor would
require at least $35$ days just to allocate one {\em machine cycle} to
every pair of nodes. This implies that in order to determine the
top-ranked link predictions over the {\em entire network}, the
running time will be much much larger than $35$ days.

@@

Latent factor models have proven a great success in the domain of
collaborative filtering, but not link prediction in spite of the obvious
similarity between link prediction and collaborative filtering and
the obvious effectiveness of latent factor models. One of the reasons why latent factor models are rarely used for
link prediction is due to their complexity. In
collaborative filtering applications, the item dimension is of the
order of a few hundred thousand, whereas even the smallest  real-world networks contain more than a million nodes.
Even worse, we also experientially find that the quality of link prediction for latent factor models decreases with the increase of data sparsity,
and networks typically become sparser when their sizes grow larger.

@@

We explore an {\em ensemble approach} to making latent factor models
practical for link prediction by decomposing the search space into a
 set of smaller matrices with three structural bagging methods with performance guarantees, which has obvious {\em
effectiveness} advantages. In this way, latent factor models only need to deal with networks with small sizes (and denser), and retain
their effectiveness and efficiency.  By incorporating with the characteristics of  link prediction, the bagging methods maintain high prediction
accuracy while reducing the network size via graph sampling techniques.
Further, the same bias-variance tradeoffs apply to the link-prediction problem, as the
standard classification problem. Therefore, the use of an ensemble
approach has obvious robustness advantages as well.

>>>>

Link prediction methods are often applied to very large and sparse networks, which have a large {\em search space} $O(n^2)$,
where $n$ is the number of nodes. Hence, the scalability is a big challenge. In fact, an often overlooked fact is that most {\em exiting link prediction algorithms evaluate the link propensities only over a subset of possibilities rather than all propensities over the entire network}. %,\eg \cite{zhu2016}.
Consider a large network with $10^8$ nodes. Its number of {\em possibilities} for links
is of the order of $10^{16}$. Therefore, a 3GHz processor would
require at least $35$ days just to allocate one {\em machine cycle} to
every pair of nodes. This implies that in order to determine the
top-ranked link predictions over the {\em entire network}, the
running time would be much more than $35$ days.

@@

Latent factor models have proven a great success for
collaborative filtering, but not link prediction in spite of the obvious
similarity and
the obvious effectiveness of latent factor models. One of the reasons why latent factor models are rarely used for
link prediction is due to their complexity. In
collaborative filtering applications, the item dimension is of the
order of a few hundred thousand, whereas even the smallest  real-world networks contain more than a million nodes.
Even worse, we also experientially find that the quality of link prediction for latent factor models decreases with the increase of data sparsity,
and networks typically become sparser when their sizes grow larger.

@@

We explore an {\em ensemble approach} to making latent factor models
practical for link prediction by decomposing the search space into a
 set of smaller matrices with three structural bagging methods with performance guarantees, which has obvious {\em
effectiveness} advantages. In this way, latent factor models only need to deal with networks with small sizes (and denser), and retain
their effectiveness and efficiency.  By incorporating with the characteristics of  link prediction, the bagging methods maintain high prediction
accuracy while reducing the network size via graph sampling techniques.
Further, the use of an ensemble
approach has obvious robustness advantages as well.

### 2018-01-27 ApproxComp/sec-introduction.tex

Over the past a few years, research and development has made significant progresses on big data analytics with the supports from both governments and industries all over the world, such as Spark\footnote{\small \url{https://spark.apache.org}}, IBM Watson\footnote{\small \url{https://www.ibm.com/watson}} and Google AlphaGo\footnote{\small \url{https://deepmind.com/research/alphago}}. A fundamental issue for big data analytics is the efficiency, and various advances towards attacking this issues have been achieved recently, from theory to algorithms to systems~\cite{FanGN13,Jordan15,ZahariaXWDADMRV16}.  In this article, we shall present the idea of two classes of techniques for efficient and effective big data analytics: query approximation and data approximation, based on our recent research experiences. Different from existing approximation techniques~\cite{CormenLRS01,approx03,FanH14}, the approximation computation that we are going to introduce is unnecessarily theoretically bounded with respect to optimal solutions, but asks for both efficiency and effectiveness in practice.

>>>>

Over the past a few years, research and development has made significant progresses on big data analytics with the supports from both governments and industries all over the world, such as Spark\footnote{\small \url{https://spark.apache.org}}, IBM Watson\footnote{\small \url{https://www.ibm.com/watson}} and Google AlphaGo\footnote{\small \url{https://deepmind.com/research/alphago}}. A fundamental issue for big data analytics is the efficiency, and various advances towards attacking this issues have been achieved recently, from theory to algorithms to systems~\cite{FanGN13,Jordan15,ZahariaXWDADMRV16}.  In this article, we present the idea of approximate computation for efficient and effective big data analytics: query approximation and data approximation, based on our recent research experiences. Different from existing approximation techniques~\cite{CormenLRS01,approx03,FanH14}, the approximation computation that we are going to introduce is unnecessarily theoretically bounded with respect to optimal solutions, but asks for both efficiency and effectiveness in practice.

### 2018-01-27 ApproxComp/sec-query.tex

\stitle{(2) One-Pass Trajectory Compression~\cite{LinMZWH17}}.  Trajectory compression (also known as trajectory simplification) is to compress data points in a trajectory to a set of continuous line segments, and is commonly used  in practice.

@@

We develop a data-driven approach, instead of filter-and-verification, to  identifying the most possible $k$ time intervals from $T \times (T + 1)/2$ time intervals, in which $T$ is the number of snapshots and k is a small constant, \eg 10. This is achieved by exploring the characteristics of time intervals involved with dense subgraphs based on the observation of {\em evolving convergence phenomenon} in traffic data, inspired by the convergent evolution in nature\footnote{\small \url{https://en.wikipedia.org/wiki/Convergent_evolution}}. That is, our method provides time intervals with probabilistic grantees, instead of exact ones with \kw{FAV}.   Using both real-life and synthetic data, we experimentally show that Our method \kw{FIDES} is over 1000 times faster than the state of the art solution  \kw{MEDEN}~\cite{BogdanovMS11}, while the quality of dense subgraphs found is comparable with  \kw{MEDEN} .

>>>>

\stitle{(2) One-Pass Trajectory Compression~\cite{LinMZWH17}}.  Trajectory compression (\aka trajectory simplification) is to compress data points in a trajectory to a set of continuous line segments, and is commonly used  in practice.

@@

We develop a data-driven approach, instead of filter-and-verification, to  identifying the most possible $k$ time intervals from $T \times (T + 1)/2$ time intervals, in which $T$ is the number of snapshots and k is a small constant, \eg 10. This is achieved by exploring the characteristics of time intervals involved with dense subgraphs based on the observation of {\em evolving convergence phenomenon} in traffic data, inspired by the convergent evolution in nature\footnote{\small \url{https://en.wikipedia.org/wiki/Convergent_evolution}}. That is, our method provides time intervals with probabilistic grantees, instead of exact ones like \kw{FAV}.   Using both real-life and synthetic data, we experimentally show that Our method \kw{FIDES} is over 1000 times faster than the state of the art solution  \kw{MEDEN}~\cite{BogdanovMS11}, while the quality of dense subgraphs found is comparable with  \kw{MEDEN} .
